---
title: "Crouse 1987 Analysis"
author: "Ellie"
date: "14 October 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
The code below demonstrates my attempt to re-create the analysis from Crouse's 1987 paper on loggerhead turtle demography. 

#Creating the Crouse 1987 Matrix

Creating a stage-based projection matrix, for each stage, 
calculating the repro- ductive output (F,), the probability of surviving and growing into the next stage (G,). and the probability of surviving and 
remaining in the same stage (P,).
This code needs the csv files 'table 3 from crouse' and 'sens'

```{r include = FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
table.3 <- read.csv("~/1 UNIVERSITY/Level 4/Project & Dissertation/Crouse 1987/table 3 from crouse.csv")
sens <- read.csv("~/1 UNIVERSITY/Level 4/Project & Dissertation/Crouse 1987/sens.csv")
fecs <- select(table.3, fecundity)
#starting values 
pi <- select(table.3, annual_survivorship)
di <-select(table.3, stage_duration)
#Ps
Pi <- ((1 - (pi^(di - 1)))/(1 - (pi^di)))*pi
# Pi's
#1              0.0000
#2              0.7371
#3              0.6611
#4              0.6907
#5              0.0000
#6              0.0000
#7              0.8086

#Gs
Gi <- (pi^di*(1 - pi))/(1 - pi^di) 
#life table (parts of anyway)
life_table <- data.frame(fecs, Gi, Pi)
##making the matrix: 
mat1 <- matrix(NA, nrow = 7, ncol = 7, byrow = T) 
#fill with fecundity 
mat1[1,] <- life_table$fecundity 
#add Ps 
mat1[1,1] <- Pi$annual_survivorship[1] 
mat1[2,2] <- Pi$annual_survivorship[2] 
mat1[3,3] <- Pi$annual_survivorship[3]
mat1[4,4] <- Pi$annual_survivorship[4]
mat1[5,5] <- Pi$annual_survivorship[5]
mat1[6,6] <- Pi$annual_survivorship[6]
mat1[7,7] <- Pi$annual_survivorship[7]
mat1
#add Gs 
mat1[2,1] <- Gi$annual_survivorship[1]
mat1[3,2] <- Gi$annual_survivorship[2]
mat1[4,3] <- Gi$annual_survivorship[3]
mat1[5,4] <- Gi$annual_survivorship[4]
mat1[6,5] <- Gi$annual_survivorship[5]
mat1[7,6] <- Gi$annual_survivorship[6] 
mat1
#remove NAs
#shows location of NAs
is.na(mat1)
#replaces NAs with 0s
mat1[is.na(mat1)] <- 0
mat1#almost there 
A <- mat1
```

**Created a function for'matrixing'**
```{r}
### making a function which creates the matrix 
lifetable <- table.3
myFunc <- function (lifetable) 
{
  fecs <- select(lifetable, fecundity)
  pi <- select(lifetable, annual_survivorship)
  di <-select(lifetable, stage_duration)
  Pi <- ((1 - (pi^(di - 1)))/(1 - (pi^di)))*pi
  Gi <- (pi^di*(1 - pi))/(1 - pi^di)
  mat1 <- matrix(0, nrow = 7, ncol = 7)
    #add Fs
  mat1[1,] <- lifetable$fecundity 
  #add Ps (diagonals)
  mat1[1,1] <- Pi$annual_survivorship[1] 
  mat1[2,2] <- Pi$annual_survivorship[2] 
  mat1[3,3] <- Pi$annual_survivorship[3]
  mat1[4,4] <- Pi$annual_survivorship[4]
  mat1[5,5] <- Pi$annual_survivorship[5]
  mat1[6,6] <- Pi$annual_survivorship[6]
  mat1[7,7] <- Pi$annual_survivorship[7]
  mat1
  #add Gs (off-diagonals)
  mat1[2,1] <- Gi$annual_survivorship[1]
  mat1[3,2] <- Gi$annual_survivorship[2]
  mat1[4,3] <- Gi$annual_survivorship[3]
  mat1[5,4] <- Gi$annual_survivorship[4]
  mat1[6,5] <- Gi$annual_survivorship[5]
  mat1[7,6] <- Gi$annual_survivorship[6] 
  return(mat1)
}

```
**Table 4** 
Stage-class population matrix for loggerhead sea turtles based on the life table presented in Table 3. For the general form of the matrix and formulae for calculating the matrix elements see Theoretical Population Projections. 
```{r}
mat1
```
**Stage structure growth (multiple steps)**
```{r include = FALSE}
N0 <- matrix(c(10000,10000,10000,10000,10000,10000,10000), ncol=1)
years <- 20
N.projections <- matrix(0, nrow = nrow(A), ncol = years + 1) 
N.projections[,1] <- N0 
N.totals <- apply(N.projections, 2, sum)
Rs <- N.totals[-1]/N.totals[-(years + 1)]
```
**Eigen analysis** 
```{r echo = FALSE} 
eigs.A <- eigen(A)
eigs.A
```
**Finding the first eigenvalue (finite rate of increase)**
```{r echo = FALSE} 
dom.pos <- which.max(eigs.A[["values"]])
L1 <- Re(eigs.A[["values"]][dom.pos])
L1
```
**Power method** 

using N0 <- matrix(c(10000,10000,10000,10000,10000,10000,10000), ncol=1) as a subsitute for actual data 
```{r}
t <- 20
Nt <- N0/sum(N0)
R.t <- numeric(t)
for (i in 1:t) R.t[i] <- {
  Nt1 <- A %*% Nt
  R <- sum(Nt1)/sum(Nt)
  R
} 
```
**Calculating the stable stage distribution** 
```{r echo = FALSE}
w <- Re(eigs.A[["vectors"]][, dom.pos])
ssd <- w / sum(w)
stable <- round(ssd,3)
stable
```
**Calculating the reproductive value** 
```{r echo = FALSE}
M <- eigen(t(A))
v <- Re(M$vectors[,which.max(Re(M$values))])
RV <- v / v[1]
RV 
```
**Created table 5 (from Crouse 1987)** 
```{r echo = FALSE}
tab <- select(table.3, stage_number, class)
tab_5<- data.frame(tab, stable, RV)
colnames(tab_5) <- c("Stage number", "Stage Class", "Stable stage distribution (Dominant eigenvector)", "Reproductive values (left eigenvector)" )
kable(tab_5, caption = "Table 5. Stable stage distribution (wJ) and reproductive values (v') for the loggerhead population matrix given in Table 4.")
```

##Sensitivity analyses
```{r echo = FALSE}
#sensitivity of projection matrices 
vw.s <- v %*% t (w) 
S <- (S <- vw.s/as.numeric(v %*% w)) 
#elasticity of projection matrices 
elas <- (A/L1) * S 
elasticity <- round(elas, 3)
### figure 3 - plot the proportional sensitivity to changes in F, P and G 
stage <- c(1:7)
F <- c(elasticity[1, 1:7])
P <- c(elasticity[1,1], elasticity[2,2], elasticity[3,3], elasticity[4,4], elasticity[5,5], elasticity[6,6], elasticity[7,7])
G <- c(elasticity[2,1], elasticity[3,2], elasticity[4,3], elasticity[5,4], elasticity[6,5], elasticity[7,6], 0)
sensitvities <- data.frame(stage, F, P, G)
fig.3 <- ggplot(sens, aes(x = stage, y = sens, colour = supp, shape = supp)) + geom_line() + geom_point(size = 4) + labs(x = "Stage", y = "Elasticity")
```
#Calculating changes in rate of increase r resulting from simulated changes in fecundity and survival of individual life history stages in the loggerhead population matrix decresing fecundity and survival by 50%
```{r echo = FALSE}
table.3_fecAdjust<-mutate(table.3, fecundity = 0.5*fecundity)
table.3_Surv1Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "1", annual_survivorship * 0.5, annual_survivorship * 1))
table.3_Surv2Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "2", annual_survivorship * 0.5, annual_survivorship * 1))
table.3_Surv3Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "3", annual_survivorship * 0.5, annual_survivorship * 1))
table.3_Surv4Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "4", annual_survivorship * 0.5, annual_survivorship * 1))
table.3_Surv5Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "5", annual_survivorship * 0.5, annual_survivorship * 1))
table.3_Surv6Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "6", annual_survivorship * 0.5, annual_survivorship * 1))
table.3_Surv7Adj <- mutate(table.3, annual_survivorship = ifelse(stage_number == "7", annual_survivorship * 0.5, annual_survivorship * 1))
#50% increase in fecundity or an increase in survivorship to 1.0.
table.3_fecInc<-mutate(table.3, fecundity = 1.5*fecundity)
table.3_Surv1Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "1", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1)) # 0.999 to test if it's a problem with 1
table.3_Surv2Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "2", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1))
table.3_Surv3Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "3", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1))
table.3_Surv4Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "4", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1))
table.3_Surv5Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "5", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1))
table.3_Surv6Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "6", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1))
table.3_Surv7Inc <- mutate(table.3, annual_survivorship = ifelse(stage_number == "7", (0.999-annual_survivorship+annual_survivorship), annual_survivorship * 1))
#eigenanalysis 
#decreases 
AFd <- myFunc(table.3_fecAdjust) 
eigs.AFd <- eigen(AFd)
A1d <- myFunc(table.3_Surv1Adj) 
eigs.A1d <- eigen(A1d)
A2d <- myFunc(table.3_Surv2Adj) 
eigs.A2d <- eigen(A2d)
A3d <- myFunc(table.3_Surv3Adj) 
eigs.A3d <- eigen(A3d)
A4d <- myFunc(table.3_Surv4Adj) 
eigs.A4d <- eigen(A4d)
A5d <- myFunc(table.3_Surv5Adj) 
eigs.A5d <- eigen(A5d)
A6d <- myFunc(table.3_Surv6Adj) 
eigs.A6d <- eigen(A6d)
A7d <- myFunc(table.3_Surv7Adj) 
eigs.A7d <- eigen(A7d)
#increases 
AFi <- myFunc(table.3_fecInc) # F for fecundity, i for increase (50%)
eigs.AFi <- eigen(AFi)
A1i <- myFunc(table.3_Surv1Inc) # 1 for stage number, i for increase (to 1.0)
eigs.A1i <- eigen(A1i)
A2i <- myFunc(table.3_Surv2Inc) 
eigs.A2i <- eigen(A2i)
A3i <- myFunc(table.3_Surv3Inc) 
eigs.A3i <- eigen(A3i)
A4i <- myFunc(table.3_Surv4Inc) 
eigs.A4i <- eigen(A4i)
A5i <- myFunc(table.3_Surv5Inc) 
eigs.A5i <- eigen(A5i)
A6i <- myFunc(table.3_Surv6Inc) 
eigs.A6i <- eigen(A6i)
A7i <- myFunc(table.3_Surv7Inc) 
eigs.A7i <- eigen(A7i)
#finding the first eigenvalue (finite rate of increase)
#decreases
dom.pos.Fd <- which.max(eigs.AFd[["values"]])
LFd <- Re(eigs.AFd[["values"]][dom.pos.Fd]) 
dom.pos.1d <- which.max(eigs.A1d[["values"]])
L1d <- Re(eigs.A1d[["values"]][dom.pos.1d]) 
dom.pos.2d <- which.max(eigs.A2d[["values"]])
L2d <- Re(eigs.A2d[["values"]][dom.pos.2d]) 
dom.pos.3d <- which.max(eigs.A3d[["values"]])
L3d <- Re(eigs.A3d[["values"]][dom.pos.3d]) 
dom.pos.4d <- which.max(eigs.A4d[["values"]])
L4d <- Re(eigs.A4d[["values"]][dom.pos.4d]) 
dom.pos.5d <- which.max(eigs.A5d[["values"]])
L5d <- Re(eigs.A5d[["values"]][dom.pos.5d]) 
dom.pos.6d <- which.max(eigs.A6d[["values"]])
L6d <- Re(eigs.A6d[["values"]][dom.pos.6d]) 
dom.pos.7d <- which.max(eigs.A7d[["values"]])
L7d <- Re(eigs.A7d[["values"]][dom.pos.7d]) 
#increases 
dom.pos.Fi <- which.max(eigs.AFi[["values"]])
LFi <- Re(eigs.AFi[["values"]][dom.pos.Fi]) 
dom.pos.1i <- which.max(eigs.A1i[["values"]])
L1i <- Re(eigs.A1i[["values"]][dom.pos.1i])
dom.pos.2i <- which.max(eigs.A2i[["values"]])
L2i <- Re(eigs.A2i[["values"]][dom.pos.2i]) 
dom.pos.3i <- which.max(eigs.A3i[["values"]])
L3i <- Re(eigs.A3i[["values"]][dom.pos.3i]) 
dom.pos.4i <- which.max(eigs.A4i[["values"]])
L4i <- Re(eigs.A4i[["values"]][dom.pos.4i]) 
dom.pos.5i <- which.max(eigs.A5i[["values"]])
L5i <- Re(eigs.A5i[["values"]][dom.pos.5i]) 
dom.pos.6i <- which.max(eigs.A6i[["values"]])
L6i <- Re(eigs.A6i[["values"]][dom.pos.6i]) 
dom.pos.7i <- which.max(eigs.A7i[["values"]])
L7i <- Re(eigs.A7i[["values"]][dom.pos.7i])
```

**Figure 1**
Changes in rate of increase r resulting from sim- ulated changes in fecundity and survival of individual life history stages in the loggerhead population matrix (remaining components held constant). The dashed line represents the r determined in the baseline run on the initial matrix. 
```{r echo = FALSE}
stage_class <- c("Fecundity", "Eggs/Hatchlings", "Small Juveniles", "Large Juveniles", "Subadults", "Novice Breeders", "1st-yr Remigrants", "Mature Breeders")
lambdas<- c(LFd, L1d, L2d, L3d, L4d, L5d, L6d, L7d)
rs <- log(lambdas)
table_decrease <- data.frame(stage_class, rs)
graph <- ggplot(table_decrease, aes(x = stage_class, y = rs)) + geom_bar(stat = "identity")
graph1 <- graph + labs(x = "Stage Class", y = "Intrinsic rate of Increase (r)") + geom_hline(aes(yintercept=-0.056399), linetype = "dashed", size = 0.75)  
graph2 <- graph1 + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())+ expand_limits(y = -0.25, x = 0.2)
figure1a <- graph2 + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + scale_x_discrete(limits = c("Fecundity", "Eggs/Hatchlings", "Small Juveniles", "Large Juveniles", "Subadults", "Novice Breeders", "1st-yr Remigrants", "Mature Breeders"))
figure1a 
stage_class <- c("Fecundity", "Eggs/Hatchlings", "Small Juveniles", "Large Juveniles", "Subadults", "Novice Breeders", "1st-yr Remigrants", "Mature Breeders")
lambdas<- c(LFi, L1i, L2i, L3i, L4i, L5i, L6i, L7i)
rs <- log(lambdas)
table_decrease <- data.frame(stage_class, rs)
graph <- ggplot(table_decrease, aes(x = stage_class, y = rs)) + geom_bar(stat = "identity")
graph1 <- graph + labs(x = "Stage Class", y = "Intrinsic rate of Increase (r)") + geom_hline(aes(yintercept=-0.056399), linetype = "dashed", size = 0.75)  
graph2 <- graph1 + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())+ expand_limits(y = -0.25, x = 0.2)
figure1b <- graph2 + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + scale_x_discrete(limits = c("Fecundity", "Eggs/Hatchlings", "Small Juveniles", "Large Juveniles", "Subadults", "Novice Breeders", "1st-yr Remigrants", "Mature Breeders"))
figure1b 
#ggplot(changes, aes(x = stage, y = rs)) + geom_bar()
```
**Figure 3** 
The elasticity, or proportional sensitivity of lambda to changes in fecundity, F, survival while remainging in the same stage (P) and survival with growth (G). Because the elasticities of this matrix sum to 1, they can be compared directly in terms of their contribution to the population growth rate r. 

```{r echo = FALSE}
fig.3
```